A Reproducible Data Analysis Workflow with RMarkdown, Git, Make & Docker

### 2019-08-23 rr-flow.Rmd
fix compile issue introduced by tabs instead of spaces
added officer package description and some limitations

Such change prevents many reports from beeing replicated, if they contain, e.g., multiple imputations, bootstrapping, simulations studies, graphics with random jitter, bayesian estimations, or similar technics which involve random sampling. However seldom does an analysis depend only on R, often a considerable number of packages is required, and they will most likely more often introduce breaking changes with updates. The whole endeavor of reproducibility is therefore at stake, every time an update is rolled out. To prevent that `docker` replicates the computational environment of analysis exactly. We do not intend to advocate that software should not be updated; the used software version should just be documented and easily replicated. Quite to the contrary with solutions like docker, it gets more comfortable than ever to safely update to new versions, by just bumping the R Version number of the docker image.

@@

@@

How to best share a reproducible workflow? We believe that, ideally, a service provider should be found that guarantuees permanent and reliable hosting of the reproducible workflow, such as the Open Science Framework (citation needed) or PsychData (citation needed). Second, to ensure that other users are legally able to benefit from the shared materials, authors must choose an appropriate license format. Typically, there is currently no single license that works for code, data, and media (such as text or figures). We encourage authors to choose appropriate license forms that do not hinder others from freely downloading, using, and moifying the shared workflows and materials while, at the same time, ensuring recognition for the time and effort invested in creating the workflow in the first place. In our experience, the Creative Commons license (CC-BY) is often appropriate for sharing texts, rmarkdown files, generated figures, and other media whereas scripts and any other computer code is often best shared according to the MIT license. Both licenses assure maximal freedom for future users while requiring the attribution of the original authors in derivative work.

>>>>

Such change prevents many reports from beeing replicated, if they contain, e.g., multiple imputations, bootstrapping, simulations studies, graphics with random jitter, Bayesian estimations using sampling algorithms (such as Markov Chain Monte Carlo), or similar technics which involve random sampling. However only rarelydoes an analysis depend only on base R, often a considerable number of packages is required, and they will most likely more often introduce breaking changes with updates. The whole endeavor of reproducibility is therefore at stake, every time an update is rolled out. To ensure long-term reproducibility, `docker` replicates the original computational environment of an analysis exactly. However, we do not intend to advocate that software should not be updated; updates typically promote bugfixes and provide new functionality; but full computational reproducibility is only achieved if the originally used software version(s) are documented. Among other things, this allows to trace back recent updates as to what change in what software caused a non-replication of the original results. Quite to the contrary with solutions like docker, it gets more comfortable than ever to safely update to new versions, by just bumping the R Version number of the docker image.

@@

## Limitations

We are aware that implementing this workflow is not straightforward and the difficulty of its implementation may vary by platform. For example, already the installation of all tools is easier on POSIX-compatible platforms such as Unix, Linux, or macOS (but not Windows). However, once such a workflow is established as a default, it can be used with minimal changes to the familiar use of the R environment. It is even possible to move the entire development process to a docker environment.

In our own experience, it is often not possible to convince all co-authors to switch to a different document processing environment, such as RMarkdown. That is, we have experienced that after writing up the first draft in Rmarkdown, one must eventually switch to a dynamically generated Word-file that is from then on used as static file that is used as basis for multiple iterations among the co-authors. This means, that ideally, one would have to manually merge the final Word file back to the RMarkdown souce, which generally is only possible manually. But we believe that even if this final step does not happen that - as long as dynamically generated elements are not manipulated - ensuring the reproducibility of the analysis and the initial draft is a big step forward. A less elegant but potentially more practical workflow may involve the officer package (CITATION NEEDED). The officer package allows the manipulation of  Word (.docx) and PowerPoint (*.pptx) documents, that is, given a formatted template file, one can dynamically paste images, tables and text into documents from R.

@@

How to best share a reproducible workflow? We believe that, ideally, a service provider should be found that guarantuees permanent and reliable hosting of the reproducible workflow, such as the Open Science Framework (citation needed) or PsychData (citation needed). Second, to ensure that other users are legally able to benefit from the shared materials, authors must choose an appropriate license format. Typically, there is currently no single license that works for code, data, and media (such as text or figures). We encourage authors to choose appropriate license forms that do not hinder others from freely downloading, using, and moifying the shared workflows and materials while, at the same time, ensuring recognition for the time and effort invested in creating the workflow in the first place. In our experience, the Creative Commons license (CC-BY) is often appropriate for sharing texts, rmarkdown files, generated figures, and other media whereas scripts and any other computer code is often best shared according to the MIT license. Both licenses assure maximal freedom for future users while requiring the attribution of the original authors in derivative work. These licenses are also in line with the recommendations by the Reproducibile Research Standard (see stodden2016enhancing).
